Population structure and phenotypic variation of \emph{Sclerotinia sclerotiorum} from dry bean (\emph{Phaseolus vulgaris}) in the United States

### 2017-11-29 doc/manuscript/manuscript.tex
update OSF citation
rerun make

All data processing and analyses were performed in a Rocker ``verse''
project container running R version 3.4.2 (Boettiger \& Eddelbuettel,
2017; R Core Team, 2017) and are openly available and reproducible at
\url{https://github.com/everhartlab/sclerotinia-366/}. Of the 16
microsatellite loci genotyped, five included compound repeats, which
made it challenging to accurately/confidently bin alleles into fragment
sizes expected for each locus based on the described repeat motif. Loci
with compound repeats were removed for the reported statistics. To
ensure the integrity of the results we additionally processed these loci
and included them in concurrent analyses. We assessed the power of our
11 markers by generating a genotype accumulation curve in the R package
\emph{poppr} version 2.5.0, looking for evidence of saturation, which
would indicate that loci were sufficiently sampled to adequately
represent the full set of haplotypes (Arnaud-Hanod et al., 2007; Kamvar
et al., 2015). To avoid including isolates potentially collected from
the same plant (which increases the probability of collecting sclerotia
from the same point of infection more than once), data were
clone-corrected on a hierarchy of Region/Source/Host/Year---meaning that
duplicated genotypes were reduced to a single observation when they were
collected in the same year from the same host cultivar located in the
same source field (wmn or producer)---for subsequent analysis. We
assessed haplotype diversity by calculating Stoddart and Taylor's index
(\emph{G}) (Stoddart \& Taylor, 1988), Shannon's index (\emph{H})
(Shannon, 1948), Simpson's index (\(\lambda\)) (Simpson, 1949), evenness
(\(E_5\)), and the expected number of multilocus haplotypes
(\emph{eMLH}) (Hurlbert, 1971; Heck et al., 1975; Pielou, 1975; Grünwald
et al., 2003). To assess the potential for random mating, we tested for
linkage disequilibrium with the index of association, \(I_A\) and its
standardized version, \(\bar{r}_d\) using 999 permutations (Brown et
al., 1980; Smith et al., 1993; Agapow \& Burt, 2001). Both haplotype
diversity and linkage disequilibrium were calculated in \emph{poppr}
(Kamvar et al., 2014).

@@

The number of observed alleles per locus ranged from two to 10 with an
average of 6.27 (Table \ref{tab:locus-stats}). Locus 20-3, which
contained only 2 alleles, showed low values of both \emph{h} (0.0533)
and evenness (0.42), indicating that there was one dominant allele
present. Analysis of the haplotype accumulation curve showed no clear
plateau for 11 or 16 loci (See section on `Loading Data and Setting
Strata' in the MLG-distribution.md\footnote{Direct link:
  \url{https://github.com/everhartlab/sclerotinia-366/blob/master/results/MLG-distribution.md\#loading-data-and-setting-strata}}
file in the supplemental files (Kamvar et al., 2017)), indicating that
we would likely obtain more multilocus haplotypes if we were to genotype
more loci.

@@

used in this study. \emph{h} = Nei's Gene Diversity (Nei, 1978). Average
\emph{h} = 0.583, average Evenness = 0.693, average no. alleles =
6.27}\tabularnewline

@@

\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
Evenness\strut

@@

\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut

@@

\subsection*{Conflicts of Interest}\label{conflicts-of-interest}
\addcontentsline{toc}{subsection}{Conflicts of Interest}

The authors declare no conflict of interest.

\subsection*{Author Contributions}\label{author-contributions}
\addcontentsline{toc}{subsection}{Author Contributions}

\begin{itemize}
\item
  \textbf{Zhian N. Kamvar} analyzed the data, contributed analysis
  tools, wrote the paper, prepared figures and tables, edited and
  reviewed drafts of the paper.
\item
  \textbf{B. Sajeewa Amaradasa} analyzed the data, contributed analysis
  tools, wrote drafts of the paper, edited and reviewed drafts of the
  paper.
\item
  \textbf{Rachana Jhala} Carried out experiments (MCG assessment,
  aggressiveness ratings, genotyping), edited and reviewed drafts of the
  paper.
\item
  \textbf{Serena McCoy} Carried out experiments (MCG assessment,
  aggressiveness ratings, genotyping), edited and reviewed drafts of the
  paper.
\item
  \textbf{James R. Steadman} Conceived and designed experiments,
  organized network of white mold screening nurseries, provided \emph{S.
  sclerotiorum} isolates, edited and reviewed drafts of the paper.
\item
  \textbf{Sydney E. Everhart} supervised data analysis, analyzed the
  data, contributed analysis tools, wrote the paper, edited and reviewed
  drafts of the paper.
\end{itemize}

@@

\hypertarget{ref-kamvar2017data}{}
Kamvar, Z. N., Amaradasa, B. S., Jhala, R., McCoy, S., Steadman, J., \&
Everhart, S. E. (2017, November). Data and analysis for population
structure and phenotypic variation of \emph{Sclerotinia sclerotiorum}
from dry bean (\emph{Phaseolus vulgaris}) in the United States. Open
Science Framework. \url{https://doi.org/10.17605/OSF.IO/EJB5Y}

\hypertarget{ref-kamvar2015novel}{}
Kamvar, Z. N., Brooks, J. C., \& Grünwald, N. J. (2015). Novel R tools
for analysis of genome-wide population genetic data with emphasis on
clonality. \emph{Frontiers in Genetics}, \emph{6}.
\url{https://doi.org/10.3389/fgene.2015.00208}

@@

\hypertarget{ref-mccoy2009use}{}
McCoy, S., \& Steadman, J. R. (2009). Use of multi-site screening to
identify partial resistance to white mold in common bean in 2008.
\emph{Bean Improvement Cooperative Annual Report}, 86--87. Retrieved
from \url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND44207142}

@@

\hypertarget{ref-mendiburu2015agricolae}{}
Mendiburu, F. D., \& Simon, R. (2015). Agricolae - ten years of an open
source statistical tool for experiments in breeding, agriculture and
biology. \url{https://doi.org/10.7287/peerj.preprints.1404v1}

@@

\hypertarget{ref-otto2007identification}{}
Otto-Hanson, L., \& Steadman, J. R. (2007). Identification of partial
resistance to \emph{Sclerotinia sclerotiorum} in common bean at multiple
locations in 2006. \emph{Bean Improvement Cooperative Annual Report},
133--134. Retrieved from
\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND43940892}

\hypertarget{ref-otto2008identification}{}
Otto-Hanson, L., \& Steadman, J. R. (2008). Identification of partial
resistance to \emph{Sclerotinia sclerotiorum} in common bean at multiple
locations in 2007. \emph{Bean Improvement Cooperative Annual Report},
214--215. Retrieved from
\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND44063230}

@@

\hypertarget{ref-steadman2003identification}{}
Steadman, J. R., Eskridge, K., \& Powers, K. (2003). Identification of
partial resistance to \emph{Sclerotinia sclerotiorum} in common bean at
multiple locations. \emph{Bean Improvement Cooperative Annual Report},
225--226. Retrieved from
\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND43757287}

\hypertarget{ref-steadman2006identification}{}
Steadman, J. R., Otto-Hanson, L., \& Breathnach, J. (2006).
Identification of partial resistance to \emph{Sclerotinia sclerotiorum}
in common bean at multiple locations in 2005. \emph{Bean Improvement
Cooperative Annual Report}, 223--224. Retrieved from
\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND43805570}

@@

\hypertarget{ref-steadman2005identification}{}
Steadman, J. R., Otto-Hanson, L., \& Powers, K. (2005). Identification
of partial resistance to \emph{Sclerotinia sclerotiorum} in common bean
at multiple locations in 2004. \emph{Bean Improvement Cooperative Annual
Report}, 124--125. Retrieved from
\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND43759243}

>>>>

All data processing and analyses were performed in a Rocker ``verse''
project container running R version 3.4.2 (Boettiger \& Eddelbuettel,
2017; R Core Team, 2017). These analyses were rendered as dynamic
documents with the R packages \emph{knitr} (version 1.17) and
\emph{ezknitr} (version 0.6) and are openly available and reproducible
at \url{https://github.com/everhartlab/sclerotinia-366/} (Attali, 2016;
Xie, 2017). Of the 16 microsatellite loci genotyped, five included
compound repeats, which made it challenging to accurately/confidently
bin alleles into fragment sizes expected for each locus based on the
described repeat motif. Loci with compound repeats were removed for the
reported statistics. To ensure the integrity of the results we
additionally processed these loci and included them in concurrent
analyses. We assessed the power of our 11 markers by generating a
genotype accumulation curve in the R package \emph{poppr} version 2.5.0,
looking for evidence of saturation, which would indicate that loci were
sufficiently sampled to adequately represent the full set of haplotypes
(Arnaud-Hanod et al., 2007; Kamvar et al., 2015). We additionally
assessed within-locus allelic diversity by measuring Nei's gene
diversity (\emph{h}) (Nei, 1978) and allelic evenness (\(E_5\)) (Pielou,
1975; Grünwald et al., 2003). To avoid including isolates potentially
collected from the same plant (which increases the probability of
collecting sclerotia from the same point of infection more than once),
data were clone-corrected on a hierarchy of
Region/Source/Host/Year---meaning that duplicated genotypes were reduced
to a single observation when they were collected in the same year from
the same host cultivar located in the same source field (wmn or
producer)---for subsequent analysis. We assessed haplotype diversity by
calculating Stoddart and Taylor's index (\emph{G}) (Stoddart \& Taylor,
1988), Shannon's index (\emph{H}) (Shannon, 1948), Simpson's index
(\(\lambda\)) (Simpson, 1949), evenness (\(E_5\)), and the expected
number of multilocus haplotypes via rarefaction (\emph{eMLH}) with 10
samples (Hurlbert, 1971; Heck et al., 1975; Pielou, 1975; Grünwald et
al., 2003). If all haplotypes in are equally abundant, both \emph{G} and
\(e^H\) (the exponentiation of \emph{H}) are expected to be equal to the
number of haplotypes, \(\lambda\) and \(E_5\) are expected to equal one,
and \emph{eMLH} is expected to be at its maximum value (in this case,
10) (Grünwald et al., 2003). To assess the potential for random mating,
we tested for linkage disequilibrium with the index of association,
\(I_A\) and its standardized version, \(\bar{r}_d\) using 999
permutations (Brown et al., 1980; Smith et al., 1993; Agapow \& Burt,
2001). Both haplotype diversity and linkage disequilibrium were
calculated in \emph{poppr} (Kamvar et al., 2014).

@@

The number of observed alleles per locus ranged from two to 10 with an
average of 6.27 (Table \ref{tab:locus-stats}). Locus 20-3, which
contained only 2 alleles, showed low values of both \emph{h} (0.0533)
and \(E_5\) (0.42), indicating that there was one dominant allele
present. Analysis of the haplotype accumulation curve showed no clear
plateau for 11 or 16 loci (See section on `Loading Data and Setting
Strata' in the MLG-distribution.md\footnote{Direct link:
  \url{https://github.com/everhartlab/sclerotinia-366/blob/master/results/MLG-distribution.md\#loading-data-and-setting-strata}}
file in the supplemental files (Kamvar et al., 2017)), indicating that
we would likely obtain more multilocus haplotypes if we were to genotype
more loci.

@@

used in this study. \emph{h} = Nei's 1978 gene diversity, \(E_5\) =
Evenness. Average \emph{h} = 0.583, average \(E_5\) = 0.693, average no.
alleles = 6.27}\tabularnewline

@@

@@

@@

@@

\hypertarget{ref-kamvar2017data}{}
Kamvar, Z. N., Amaradasa, B. S., Jhala, R., McCoy, S., Steadman, J. R.,
\& Everhart, S. E. (2017, November). Data and analysis for population
structure and phenotypic variation of \emph{Sclerotinia sclerotiorum}
from dry bean (\emph{Phaseolus vulgaris}) in the United States. Open
Science Framework. \url{https://doi.org/10.17605/OSF.IO/K8WTM}

\hypertarget{ref-kamvar2015novel}{}
Kamvar, Z. N., Brooks, J. C., \& Grünwald, N. J. (2015). Novel R tools
for analysis of genome-wide population genetic data with emphasis on
clonality. \emph{Frontiers in Genetics}, \emph{6}, 208.
\url{https://doi.org/10.3389/fgene.2015.00208}

@@

\hypertarget{ref-mccoy2009use}{}
McCoy, S., \& Steadman, J. R. (2009). Use of multi-site screening to
identify partial resistance to white mold in common bean in 2008.
\emph{Bean Improvement Cooperative Annual Report}, \emph{52}, 86--87.
Retrieved from
\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND44207142}

@@

\hypertarget{ref-mendiburu2015agricolae}{}
Mendiburu, F. D., \& Simon, R. (2015). Agricolae - ten years of an open
source statistical tool for experiments in breeding, agriculture and
biology. \emph{PeerJ PrePrints}, \emph{3}, e1404v1.
\url{https://doi.org/10.7287/peerj.preprints.1404v1}

@@

\hypertarget{ref-otto2007identification}{}
Otto-Hanson, L., \& Steadman, J. R. (2007). Identification of partial
resistance to \emph{Sclerotinia sclerotiorum} in common bean at multiple
locations in 2006. \emph{Bean Improvement Cooperative Annual Report},
\emph{50}, 133--134. Retrieved from
\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND43940892}

\hypertarget{ref-otto2008identification}{}
Otto-Hanson, L., \& Steadman, J. R. (2008). Identification of partial
resistance to \emph{Sclerotinia sclerotiorum} in common bean at multiple
locations in 2007. \emph{Bean Improvement Cooperative Annual Report},
\emph{51}, 214--215. Retrieved from
\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND44063230}

@@

\hypertarget{ref-steadman2003identification}{}
Steadman, J. R., Eskridge, K., \& Powers, K. (2003). Identification of
partial resistance to \emph{Sclerotinia sclerotiorum} in common bean at
multiple locations. \emph{Bean Improvement Cooperative Annual Report},
\emph{46}, 225--226. Retrieved from
\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND43757287}

\hypertarget{ref-steadman2006identification}{}
Steadman, J. R., Otto-Hanson, L., \& Breathnach, J. (2006).
Identification of partial resistance to \emph{Sclerotinia sclerotiorum}
in common bean at multiple locations in 2005. \emph{Bean Improvement
Cooperative Annual Report}, \emph{49}, 223--224. Retrieved from
\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND43805570}

@@

\hypertarget{ref-steadman2005identification}{}
Steadman, J. R., Otto-Hanson, L., \& Powers, K. (2005). Identification
of partial resistance to \emph{Sclerotinia sclerotiorum} in common bean
at multiple locations in 2004. \emph{Bean Improvement Cooperative Annual
Report}, \emph{48}, 124--125. Retrieved from
\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND43759243}

### 2017-11-29 doc/manuscript/paper.tex
update citations
rerun make

All data processing and analyses were performed in a Rocker ``verse''
project container running R version 3.4.2 (Boettiger \& Eddelbuettel,
2017; R Core Team, 2017). These analyses were rendered as dynamic
documents with the R packages \emph{knitr} (version 1.17) and
\emph{ezknitr} (version 0.6) and are openly available and reproducible
at \url{https://github.com/everhartlab/sclerotinia-366/} (Attali, 2016;
Xie, 2017). Of the 16 microsatellite loci genotyped, five included
compound repeats, which made it challenging to accurately/confidently
bin alleles into fragment sizes expected for each locus based on the
described repeat motif. Loci with compound repeats were removed for the
reported statistics. To ensure the integrity of the results we
additionally processed these loci and included them in concurrent
analyses. We assessed the power of our 11 markers by generating a
genotype accumulation curve in the R package \emph{poppr} version 2.5.0,
looking for evidence of saturation, which would indicate that loci were
sufficiently sampled to adequately represent the full set of haplotypes
(Arnaud-Hanod et al., 2007; Kamvar et al., 2015). To avoid including
isolates potentially collected from the same plant (which increases the
probability of collecting sclerotia from the same point of infection
more than once), data were clone-corrected on a hierarchy of
Region/Source/Host/Year---meaning that duplicated genotypes were reduced
to a single observation when they were collected in the same year from
the same host cultivar located in the same source field (wmn or
producer)---for subsequent analysis. We assessed haplotype diversity by
calculating Stoddart and Taylor's index (\emph{G}) (Stoddart \& Taylor,
1988), Shannon's index (\emph{H}) (Shannon, 1948), Simpson's index
(\(\lambda\)) (Simpson, 1949), evenness (\(E_5\)), and the expected
number of multilocus haplotypes (\emph{eMLH}) (Hurlbert, 1971; Heck et
al., 1975; Pielou, 1975; Grünwald et al., 2003). To assess the potential
for random mating, we tested for linkage disequilibrium with the index
of association, \(I_A\) and its standardized version, \(\bar{r}_d\)
using 999 permutations (Brown et al., 1980; Smith et al., 1993; Agapow
\& Burt, 2001). Both haplotype diversity and linkage disequilibrium were
calculated in \emph{poppr} (Kamvar et al., 2014).

@@

\hypertarget{ref-kamvar2015novel}{}
Kamvar, Z. N., Brooks, J. C., \& Grünwald, N. J. (2015). Novel R tools
for analysis of genome-wide population genetic data with emphasis on
clonality. \emph{Frontiers in Genetics}, \emph{6}.
\url{https://doi.org/10.3389/fgene.2015.00208}

>>>>

All data processing and analyses were performed in a Rocker ``verse''
project container running R version 3.4.2 (Boettiger \& Eddelbuettel,
2017; R Core Team, 2017). These analyses were rendered as dynamic
documents with the R packages \emph{knitr} (version 1.17) and
\emph{ezknitr} (version 0.6) and are openly available and reproducible
at \url{https://github.com/everhartlab/sclerotinia-366/} (Attali, 2016;
Xie, 2017). Of the 16 microsatellite loci genotyped, five included
compound repeats, which made it challenging to accurately/confidently
bin alleles into fragment sizes expected for each locus based on the
described repeat motif. Loci with compound repeats were removed for the
reported statistics. To ensure the integrity of the results we
additionally processed these loci and included them in concurrent
analyses. We assessed the power of our 11 markers by generating a
genotype accumulation curve in the R package \emph{poppr} version 2.5.0,
looking for evidence of saturation, which would indicate that loci were
sufficiently sampled to adequately represent the full set of haplotypes
(Arnaud-Hanod et al., 2007; Kamvar et al., 2015). We additionally
assessed within-locus allelic diversity by measuring Nei's gene
diversity (\emph{h}) (Nei, 1978) and allelic evenness (\(E_5\)) (Pielou,
1975; Grünwald et al., 2003). To avoid including isolates potentially
collected from the same plant (which increases the probability of
collecting sclerotia from the same point of infection more than once),
data were clone-corrected on a hierarchy of
Region/Source/Host/Year---meaning that duplicated genotypes were reduced
to a single observation when they were collected in the same year from
the same host cultivar located in the same source field (wmn or
producer)---for subsequent analysis. We assessed haplotype diversity by
calculating Stoddart and Taylor's index (\emph{G}) (Stoddart \& Taylor,
1988), Shannon's index (\emph{H}) (Shannon, 1948), Simpson's index
(\(\lambda\)) (Simpson, 1949), evenness (\(E_5\)), and the expected
number of multilocus haplotypes via rarefaction (\emph{eMLH}) with 10
samples (Hurlbert, 1971; Heck et al., 1975; Pielou, 1975; Grünwald et
al., 2003). If all haplotypes in are equally abundant, both \emph{G} and
\(e^H\) (the exponentiation of \emph{H}) are expected to be equal to the
number of haplotypes, \(\lambda\) and \(E_5\) are expected to equal one,
and \emph{eMLH} is expected to be at its maximum value (in this case,
10) (Grünwald et al., 2003). To assess the potential for random mating,
we tested for linkage disequilibrium with the index of association,
\(I_A\) and its standardized version, \(\bar{r}_d\) using 999
permutations (Brown et al., 1980; Smith et al., 1993; Agapow \& Burt,
2001). Both haplotype diversity and linkage disequilibrium were
calculated in \emph{poppr} (Kamvar et al., 2014).

@@

\hypertarget{ref-kamvar2015novel}{}
Kamvar, Z. N., Brooks, J. C., \& Grünwald, N. J. (2015). Novel R tools
for analysis of genome-wide population genetic data with emphasis on
clonality. \emph{Frontiers in Genetics}, \emph{6}, 208.
\url{https://doi.org/10.3389/fgene.2015.00208}
